\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-01-05 redaction/dhp_july2017.tex
relecture_JH
relecture modifs Guillaume

\title{\textsc{Transport costs did decline since 1974:\\Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
 In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. This difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Exploiting information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel), three main results emerge. First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively. Second, taking additive costs into account significantly improves the fit of the modeling of transport costs. Third, the modeling of the additive component is key to computing  the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs for given goods and origins.

@@

Third, additive costs matter in accounting for the time trend of transport costs, as we show by exploiting the time dimension of our database to document the patterns of transport costs over time. Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components. Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the modeling of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip

>>>>

\title{\textsc{Transport costs did decline since 1974:\\Findings from an explicit consideration of additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
 In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. This difference of results can be attributed to the new way of accounting for the additive component of transport costs we offer. Exploiting information contained in the US imports flows database over 1974-2013, distinguishing by transport mode (air and vessel), three main results emerge. First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively. Second, taking additive costs into account significantly improves the fit of the modeling of transport costs. Third, the modeling of the additive component is key to computing the decomposition of the trend patterns of transport costs, between what is attributable to trade composition effects and changes in the transport costs for given goods and origins.

@@

Third, additive costs matter in accounting for the time trend of transport costs, as we show by exploiting the time dimension of our database to document the patterns of transport costs over time. Between 1974 and 2013, we thus find that transport costs have decreased by roughly 50\% in air shipping and 60\% in vessel shipping, and of same order of magnitude for both additive and iceberg components. Importantly, we find that most of this decline can be attributed to a reduction in the ``ceteris paribus'' transport costs, rather than to trade composition effects, in contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). We also show that the difference of results can be attributed to the explicit consideration of the additive component: Allowing the additive costs to vary over the triple time/product/country partner case (rather than being constant as in \cite{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. In all three aspects, our results thus point the importance of the additive component in accounting for international transport costs.\smallskip
